\TitlePt

### 2016-03-01 paper1.tex
Add "p1" to labels to avoid name clashes with other papers

The \texttt{fatiando.vis.mpl} module imports all the functions in
\texttt{matplotlib.pyplot}, adds new functions, and overwrites others to
automate repetitive tasks (such as griding). Thus, the basic
functionality of the \texttt{pyplot} interface is maintained while
customizations facilitate common tasks. The following example
illustrates the use of the custom \texttt{fatiando.vis.mpl.contourf}
function to automatically grid and plot some irregularly sampled data
(Figure~\ref{fig:contourf}):

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/paper1/gridding_plotting_contourf}
    \caption{
    Example of 1) generating a random scatter of points (black dots),
    2) using that to make synthetic data, and
    3) automatically gridding and plotting the data using a
    Fatiando a Terra wrapper for the Matplotlib ``contourf``
    function.
    }
    \label{fig:contourf}
\end{figure}

@@

Map projections in Matplotlib are handled by the
\href{http://matplotlib.org/basemap}{Basemap toolkit}. The
\texttt{fatiando.vis.mpl} module also provides helper functions to
automate the use of this toolkit. The \texttt{fatiando.vis.mpl.basemap}
function automates the creation of the \texttt{Basemap} objects with
common parameters. This object can then be passed to the
\texttt{contourf}, \texttt{contour} and \texttt{pcolor} functions in
\texttt{fatiando.vis.mpl} and they will automatically plot using the
given projection (Figure~\ref{fig:basemap}):

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/paper1/gridding_plotting_basemap}
    \caption{
        Example of map plotting with the Robinson projection using the
        Matplotlib Basemap toolkit.
    }
    \label{fig:basemap}
\end{figure}

@@

The \texttt{fatiando.vis.myv} module contains functions to automate 3D plotting
using Mayavi \citep{ramachandran2011}. The \texttt{mayavi.mlab} interface
requires geometric elements to be formatted as TVTK objects. Thus, plotting
functions in \texttt{fatiando.vis.myv} automatically create TVTK
representations of \texttt{fatiando.mesher} objects and plot them using a
suitable function of \texttt{mayavi.mlab}. Also included are utility functions
for drawing axes, walls on the figure bounding box, etc. For example, the
\texttt{fatiando.vis.myv.figure} function creates a figure and rotates it so
that the z-axis points down, as is standard in geophysics. The following
example shows how to plot the 3D right rectangular prism model that we created
previously (Figure~\ref{fig:twoprisms}):

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/meshes_3dplotting_2prisms}
    \caption{
        Example of plotting a list of right rectangular prisms in Mayavi.
    }
    \label{fig:twoprisms}
\end{figure}

The \texttt{fatiando.mesher} module also contains classes for
collections of elements (e.g., meshes). A good example is the
\texttt{PrismMesh} class that represents a structured mesh of right
rectangular prisms. This class behaves as a list of
\texttt{fatiando.mesher.Prism} objects and can be passed to functions
that ask for a list of prisms, like \texttt{fatiando.vis.myv.prisms}.
Physical properties can be assigned to the mesh using the
\texttt{addprop} method (Figure~\ref{fig:mesh}):

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/meshes_3dplotting_mesh}
    \caption{
        Example of generating and visualizing a structured prism mesh.
    }
    \label{fig:mesh}
\end{figure}

Often times the mesh is used to make a detailed model of an irregular
region of the Earth's surface. In such cases, it is necessary to
consider the topography of the region. The \texttt{PrismMesh} class has
a \texttt{carvetopo} method that masks the prisms that fall above the
topography. The example below illustrates this functionality using
synthetic topography (Figure~\ref{fig:meshtopo}):

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/meshes_3dplotting_meshtopo}
    \caption{
        Example of generating and visualizing a prism mesh with masked
        topography.
    }
    \label{fig:meshtopo}
\end{figure}

When modeling involves the whole Earth, or a large area of it, the
geophysicist needs to take into account the Earth's curvature. In such
cases, rectangular prisms are inadequate for modeling and tesseroids
(e.g., spherical prisms) are better suited. The
\texttt{fatiando.vis.myv} module contains auxiliary functions to plot
along with tesseroids: an Earth-sized sphere, meridians and parallels,
as well as continental borders (Figure~\ref{fig:tesseroid}):

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/meshes_3dplotting_tesseroid}
    \caption{
        Example of creating a tesseroid (spherical prism) model and visualizing
        it in Mayavi.
    }
    \label{fig:tesseroid}
\end{figure}

@@

The Fatiando a Terra packages have separate modules for forward modeling
and inversion algorithms. The forward modeling functions usually take as
arguments geometric elements from \texttt{fatiando.mesher} with assigned
physical properties and return the synthetic data. For example, the
module \texttt{fatiando.gravmag.tesseroid} is a Python implementation of
the program Tesseroids (\url{http://leouieda.github.io/tesseroids}) and
calculates the gravitational fields of tesseroids (e.g., spherical
prisms). The following example shows how to calculate the gravity
anomaly of the tesseroid model generated in the previous section
(Figure~\ref{fig:tesseroidgrav}):

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/gravmag_tesseroid_data}
    \caption{
        Example of forward modeling the gravity anomaly using the tesseroid
        model shown in Figure~\ref{fig:tesseroid}.
    }
    \label{fig:tesseroidgrav}
\end{figure}

The module \texttt{fatiando.gravmag.polyprism} implements the method of
\citet{plouff1976} to forward model the gravity fields of a 3D right polygonal
prism. The following code sample shows how to interactively generate a
polygonal prism model and calculate its gravity anomaly
(Figures~\ref{fig:drawing} and \ref{fig:polyprism}):

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/forward_modeling_polyprism_drawing}
    \caption{
        Screen-shot of interactively drawing the contour of a 3D polygonal
        prism, as viewed from above.
    }
    \label{fig:drawing}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/paper1/forward_modeling_polyprism}
    \caption{
        Example of forward modeling the gravity anomaly of a 3D polygonal
        prism.
        a) forward modeled gravity anomaly.
        b) 3D plot of the polygonal prism.
    }
    \label{fig:polyprism}
\end{figure}

@@

The module \texttt{fatiando.gravmag.imaging} implements the imaging methods
described in \citet{fedi2012}. These methods aim to produce an image of the
geologic source from the observed gravity or magnetic data. The following code
sample uses the ``sandwich model'' method \citep{pedersen1991} to image the
polygonal prism, produced in the previous section, based on its gravity anomaly
(Figure~\ref{fig:imaging}):

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/gravmag_imaging}
    \caption{
        Example of using the "sandwich model" imaging method to recover a 3D
        image of a geologic body based on its gravity anomaly. The colored
        blocks are a cutoff of the imaged body. The black contours are the true
        source of the gravity anomaly.
    }
    \label{fig:imaging}
\end{figure}

Also implemented in Fatiando a Terra are some recent developments in
gravity and magnetic inversion methods. The method of ``planting
anomalous densities'' by \citet{uieda2012} is implemented in the
\texttt{fatiando.gravmag.harvester} module. In contrast to imaging
methods, this is an inversion method, i.e., it estimates a physical
property distribution (density in the case of gravity data) that fits
the observed data. This particular method requires the user to specify a
``seed'' (Figure~\ref{fig:seed}) around which the estimated density
distribution grows (Figure~\ref{fig:harvester}):

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/gravmag_harvester_seed}
    \caption{
        The small blue prism is the seed used by
        \texttt{fatiando.gravmag.harvester} to perform the inversion of a
        gravity anomaly. The black contours are the true source of the gravity
        anomaly.
    }
    \label{fig:seed}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/gravmag_harvester}
    \caption{
        The blue prisms are the result of a gravity inversion using module
        \texttt{fatiando.gravmag.harvester}. The black contours are the true
        source of the gravity anomaly. Notice how the inversion was able to
        recover the approximate geometry of the true source.
    }
    \label{fig:harvester}
\end{figure}

@@

The following example uses module \texttt{fatiando.seismic.srtomo} to
perform a simplified 2D tomography on synthetic seismic wave travel-time
data. To generate the travel-times we used a seismic wave velocity model
constructed from an image file. The colors of the image are converted to
gray-scale and the intensity is mapped to seismic wave velocity by the
\texttt{img2prop} method of the \texttt{fatiando.mesher.SquareMesh}
class. This model (Figure~\ref{fig:tomo}) is then used to calculate the
travel-times between a random set of earthquake locations and seismic
receivers (seismometers):

@@

Even though the implementation in \texttt{fatiando.seismic.srtomo} is greatly
simplified and not usable in real tomography problems, the result in
Figure~\ref{fig:tomo} illustrates interesting inverse problem concepts.  Notice
how the estimated velocity is blurred in the corners where no rays pass
through. This is because the data (travel-times) provide no information about
the velocity in those areas. Areas like those constitute the null space of the
inverse problem \citep{menke1984}, where any velocity value estimated will
provide an equal fit to the data.  Thus, the tomography problem requires the
use of prior information in the form of regularization. Most commonly used in
tomography problems is the Tikhonov first-order regularization, e.g., a
smoothness constraint \citep{menke1984}. The amount of smoothness imposed on
the solution is controlled by the \texttt{smooth} argument of function
\texttt{fatiando.seismic.srtomo.run}. That is how we are able to estimate a
unique and stable solution and why the result is specially smoothed where there
are no rays.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/paper1/seismic_tomo}
    \caption{
        Example run of a simplified 2D tomography. The top-left panel shows the
        true velocity model with the locations of earthquakes (yellow stars)
        and receivers (green triangles). The top-right panel shows the
        ray-paths between earthquakes and receivers. The bottom-left panel is
        the velocity estimated by the tomography. The bottom-right panel is a
        histogram of the travel-time residuals of the tomography. Notice how
        the majority of residuals are close to 0 s, indicating a good fit to
        the data.
    }
    \label{fig:tomo}
\end{figure}

>>>>

The \texttt{fatiando.vis.mpl} module imports all the functions in
\texttt{matplotlib.pyplot}, adds new functions, and overwrites others to
automate repetitive tasks (such as griding). Thus, the basic
functionality of the \texttt{pyplot} interface is maintained while
customizations facilitate common tasks. The following example
illustrates the use of the custom \texttt{fatiando.vis.mpl.contourf}
function to automatically grid and plot some irregularly sampled data
(Figure~\ref{fig:p1-contourf}):

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/paper1/gridding_plotting_contourf}
    \caption{
    Example of 1) generating a random scatter of points (black dots),
    2) using that to make synthetic data, and
    3) automatically gridding and plotting the data using a
    Fatiando a Terra wrapper for the Matplotlib ``contourf``
    function.
    }
    \label{fig:p1-contourf}
\end{figure}

@@

Map projections in Matplotlib are handled by the
\href{http://matplotlib.org/basemap}{Basemap toolkit}. The
\texttt{fatiando.vis.mpl} module also provides helper functions to
automate the use of this toolkit. The \texttt{fatiando.vis.mpl.basemap}
function automates the creation of the \texttt{Basemap} objects with
common parameters. This object can then be passed to the
\texttt{contourf}, \texttt{contour} and \texttt{pcolor} functions in
\texttt{fatiando.vis.mpl} and they will automatically plot using the
given projection (Figure~\ref{fig:p1-basemap}):

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/paper1/gridding_plotting_basemap}
    \caption{
        Example of map plotting with the Robinson projection using the
        Matplotlib Basemap toolkit.
    }
    \label{fig:p1-basemap}
\end{figure}

@@

The \texttt{fatiando.vis.myv} module contains functions to automate 3D plotting
using Mayavi \citep{ramachandran2011}. The \texttt{mayavi.mlab} interface
requires geometric elements to be formatted as TVTK objects. Thus, plotting
functions in \texttt{fatiando.vis.myv} automatically create TVTK
representations of \texttt{fatiando.mesher} objects and plot them using a
suitable function of \texttt{mayavi.mlab}. Also included are utility functions
for drawing axes, walls on the figure bounding box, etc. For example, the
\texttt{fatiando.vis.myv.figure} function creates a figure and rotates it so
that the z-axis points down, as is standard in geophysics. The following
example shows how to plot the 3D right rectangular prism model that we created
previously (Figure~\ref{fig:p1-twoprisms}):

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/meshes_3dplotting_2prisms}
    \caption{
        Example of plotting a list of right rectangular prisms in Mayavi.
    }
    \label{fig:p1-twoprisms}
\end{figure}

The \texttt{fatiando.mesher} module also contains classes for
collections of elements (e.g., meshes). A good example is the
\texttt{PrismMesh} class that represents a structured mesh of right
rectangular prisms. This class behaves as a list of
\texttt{fatiando.mesher.Prism} objects and can be passed to functions
that ask for a list of prisms, like \texttt{fatiando.vis.myv.prisms}.
Physical properties can be assigned to the mesh using the
\texttt{addprop} method (Figure~\ref{fig:p1-mesh}):

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/meshes_3dplotting_mesh}
    \caption{
        Example of generating and visualizing a structured prism mesh.
    }
    \label{fig:p1-mesh}
\end{figure}

Often times the mesh is used to make a detailed model of an irregular
region of the Earth's surface. In such cases, it is necessary to
consider the topography of the region. The \texttt{PrismMesh} class has
a \texttt{carvetopo} method that masks the prisms that fall above the
topography. The example below illustrates this functionality using
synthetic topography (Figure~\ref{fig:p1-meshtopo}):

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/meshes_3dplotting_meshtopo}
    \caption{
        Example of generating and visualizing a prism mesh with masked
        topography.
    }
    \label{fig:p1-meshtopo}
\end{figure}

When modeling involves the whole Earth, or a large area of it, the
geophysicist needs to take into account the Earth's curvature. In such
cases, rectangular prisms are inadequate for modeling and tesseroids
(e.g., spherical prisms) are better suited. The
\texttt{fatiando.vis.myv} module contains auxiliary functions to plot
along with tesseroids: an Earth-sized sphere, meridians and parallels,
as well as continental borders (Figure~\ref{fig:p1-tesseroid}):

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/meshes_3dplotting_tesseroid}
    \caption{
        Example of creating a tesseroid (spherical prism) model and visualizing
        it in Mayavi.
    }
    \label{fig:p1-tesseroid}
\end{figure}

@@

The Fatiando a Terra packages have separate modules for forward modeling
and inversion algorithms. The forward modeling functions usually take as
arguments geometric elements from \texttt{fatiando.mesher} with assigned
physical properties and return the synthetic data. For example, the
module \texttt{fatiando.gravmag.tesseroid} is a Python implementation of
the program Tesseroids (\url{http://leouieda.github.io/tesseroids}) and
calculates the gravitational fields of tesseroids (e.g., spherical
prisms). The following example shows how to calculate the gravity
anomaly of the tesseroid model generated in the previous section
(Figure~\ref{fig:p1-tesseroidgrav}):

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/gravmag_tesseroid_data}
    \caption{
        Example of forward modeling the gravity anomaly using the tesseroid
        model shown in Figure~\ref{fig:p1-tesseroid}.
    }
    \label{fig:p1-tesseroidgrav}
\end{figure}

The module \texttt{fatiando.gravmag.polyprism} implements the method of
\citet{plouff1976} to forward model the gravity fields of a 3D right polygonal
prism. The following code sample shows how to interactively generate a
polygonal prism model and calculate its gravity anomaly
(Figures~\ref{fig:p1-drawing} and \ref{fig:p1-polyprism}):

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/forward_modeling_polyprism_drawing}
    \caption{
        Screen-shot of interactively drawing the contour of a 3D polygonal
        prism, as viewed from above.
    }
    \label{fig:p1-drawing}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/paper1/forward_modeling_polyprism}
    \caption{
        Example of forward modeling the gravity anomaly of a 3D polygonal
        prism.
        a) forward modeled gravity anomaly.
        b) 3D plot of the polygonal prism.
    }
    \label{fig:p1-polyprism}
\end{figure}

@@

The module \texttt{fatiando.gravmag.imaging} implements the imaging methods
described in \citet{fedi2012}. These methods aim to produce an image of the
geologic source from the observed gravity or magnetic data. The following code
sample uses the ``sandwich model'' method \citep{pedersen1991} to image the
polygonal prism, produced in the previous section, based on its gravity anomaly
(Figure~\ref{fig:p1-imaging}):

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/gravmag_imaging}
    \caption{
        Example of using the "sandwich model" imaging method to recover a 3D
        image of a geologic body based on its gravity anomaly. The colored
        blocks are a cutoff of the imaged body. The black contours are the true
        source of the gravity anomaly.
    }
    \label{fig:p1-imaging}
\end{figure}

Also implemented in Fatiando a Terra are some recent developments in
gravity and magnetic inversion methods. The method of ``planting
anomalous densities'' by \citet{uieda2012} is implemented in the
\texttt{fatiando.gravmag.harvester} module. In contrast to imaging
methods, this is an inversion method, i.e., it estimates a physical
property distribution (density in the case of gravity data) that fits
the observed data. This particular method requires the user to specify a
``seed'' (Figure~\ref{fig:p1-seed}) around which the estimated density
distribution grows (Figure~\ref{fig:p1-harvester}):

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/gravmag_harvester_seed}
    \caption{
        The small blue prism is the seed used by
        \texttt{fatiando.gravmag.harvester} to perform the inversion of a
        gravity anomaly. The black contours are the true source of the gravity
        anomaly.
    }
    \label{fig:p1-seed}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/paper1/gravmag_harvester}
    \caption{
        The blue prisms are the result of a gravity inversion using module
        \texttt{fatiando.gravmag.harvester}. The black contours are the true
        source of the gravity anomaly. Notice how the inversion was able to
        recover the approximate geometry of the true source.
    }
    \label{fig:p1-harvester}
\end{figure}

@@

The following example uses module \texttt{fatiando.seismic.srtomo} to
perform a simplified 2D tomography on synthetic seismic wave travel-time
data. To generate the travel-times we used a seismic wave velocity model
constructed from an image file. The colors of the image are converted to
gray-scale and the intensity is mapped to seismic wave velocity by the
\texttt{img2prop} method of the \texttt{fatiando.mesher.SquareMesh}
class. This model (Figure~\ref{fig:p1-tomo}) is then used to calculate the
travel-times between a random set of earthquake locations and seismic
receivers (seismometers):

@@

Even though the implementation in \texttt{fatiando.seismic.srtomo} is greatly
simplified and not usable in real tomography problems, the result in
Figure~\ref{fig:p1-tomo} illustrates interesting inverse problem concepts.  Notice
how the estimated velocity is blurred in the corners where no rays pass
through. This is because the data (travel-times) provide no information about
the velocity in those areas. Areas like those constitute the null space of the
inverse problem \citep{menke1984}, where any velocity value estimated will
provide an equal fit to the data.  Thus, the tomography problem requires the
use of prior information in the form of regularization. Most commonly used in
tomography problems is the Tikhonov first-order regularization, e.g., a
smoothness constraint \citep{menke1984}. The amount of smoothness imposed on
the solution is controlled by the \texttt{smooth} argument of function
\texttt{fatiando.seismic.srtomo.run}. That is how we are able to estimate a
unique and stable solution and why the result is specially smoothed where there
are no rays.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/paper1/seismic_tomo}
    \caption{
        Example run of a simplified 2D tomography. The top-left panel shows the
        true velocity model with the locations of earthquakes (yellow stars)
        and receivers (green triangles). The top-right panel shows the
        ray-paths between earthquakes and receivers. The bottom-left panel is
        the velocity estimated by the tomography. The bottom-right panel is a
        histogram of the travel-time residuals of the tomography. Notice how
        the majority of residuals are close to 0 s, indicating a good fit to
        the data.
    }
    \label{fig:p1-tomo}
\end{figure}
